Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-04-17 apj.tex
brian comment

We find weak correlations between our color-excess parameters and the input features, the 
\color{orange}
only of significance much larger than $1 \sigma$
\color{black}
being
%---
$-0.17$
%---
 between
$E_{\gamma^0}(B-V)$ and $EW_{Si}$.
Recall that $EW_{Si}$ is correlated with light-curve shape, which is correlated with host-galaxy (including dust) properties 
\citep{2000AJ....120.1479H, 2003MNRAS.340.1057S}.
An extensive discussion of the correlations between the spectral features of the SNfactory data set can be found in \citet{chotard:thesis}
and \citet{leget:thesis}.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/sn2014j.pdf} 
   \caption{Top: Measured color excess for SN~2014J and the best-fit predictions from this article.   Bottom: Residual color excess
   between the data and model, with a line at zero shown for reference.
   This can be compared with the best-fit results of
   \citet{2014ApJ...788L..21A} shown in their Figure~3.
   \label{sn2014j:fig}}
\end{figure}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=5.2in]{output23/deltaratio.pdf}
   \caption{Significance of the non-zero values of $\delta$ as demonstrated by the best-fit values of $\delta_X/\delta_I-1$ for $\delta$ for Model~II.
   A line at zero is shown for reference.
   \label{deltaratio:fig}}
\end{figure}

@@

\color{black}
The analysis is validated using simulated datasets with and without third color parameters: the input parameters are recovered within  uncertainties.
We present here the results of one of the validation tests.  The input values of the supernova parameters are the median values from the analysis of real data.
Ten realizations of simulated data are generated based on the data covariances, and then analyzed separately.   The posteriors of the ten realizations
are aggregated and shown in Figures~\ref{simglobal1:fig}--\ref{simglobal5:fig}.  The input parameters are recovered with negligible bias
relative to the statistical uncertainties.  There is a tendency for the slope parameters to be slightly underestimated, which is an expected consequence of 
the uncertainties in the predictor variable, which in turn lead to 
regression dilution.  Our analysis is susceptible to this bias because we consciously do not include distributions
of supernova parameters as part of our model.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output23/residualx1.pdf}
    \caption{Differences between observed colors and the colors predicted from the analysis, as a function
            of the SALT2 light-curve shape parameter $x_1$.  For reference a dotted line is plotted at zero difference.
   \label{x1res:fig}}
\end{figure}

@@

\citet{2009ApJ...699L.139W, 2011ApJ...729...55F} find a connection between $v_{Si}$ and color, and  
\citet{2015MNRAS.447.1247S} show that $v_{Si}$ is an important spectral classifier within the SNfactory data themselves.
To first order, the $\lambda_{Si}$ treated in this article varies linearly with $v_{Si}$.
The values of $\eta$ in this article are consistent with zero.  $EW_{Ca}$ and $EW_{Si}$ have a significant effect on color,
and in turn $\lambda_{Si}$ is correlated with $EW_{Ca}$.
Removing from our model the dependence on equivalent widths (eliminating the  $\alpha$ and $\beta$ parameters), we recover
non-zero $\eta$ values at  $\gtrsim 2\sigma$.
The Ca and Si responsible for the equivalent widths and wavelength of our selected spectral features
are expected to be produced together in the thermonuclear burning into intermediate mass elements.
We conclude that $v_{Si}$ is correlated with color, 
but this correlation
is absorbed into the equivalent-width corrections, which give higher likelihood as they give lower residual dispersions $C_c$.

@@

Although the $D$ parameter has ben referred to as ``intrinsic'', due to its placement in the model and the correspondence
of the $k$ parameters with those of extrinsic two-parameter dust models, the analysis cannot quantitatively distinguish
whether its origin is intrinsic or due to a third dust parameter or some other extrinsic effect.

@@

\subsection{Summary}
To summarize, we model SNe~Ia broadband optical peak magnitudes allowing for correlations with spectral features at peak and
extrinsic color parameters.  Analyzing SNfactory data with this model, we find significant evidence that the above parameters do
affect supernova magnitudes and colors.  This model  does an excellent job in
describing SN~2014J, a supernova with extreme colors that was not used in the training.  
This work represents the first determination of the dust extinction curve outside the Local Group
derived entirely independently of assumptions about the shape of the extinction curve and/or assumptions about the
distribution of $A_V$.  An extended version
of the model retains the features of the first model, and gives the first identification of  a new parameter that affects supernova
colors in a manner that is distinct from  the expectations from dust or the spectral features $EW_{Ca}$, $EW_{Si}$, and $\lambda_{Si}$.

An interesting direction for future work is to use narrower bands for generating synthetic photometry.  When the bandwidth were small enough
to resolve spectral features, the analysis would produce a spectroscopic model.   Higher resolution would allow us to forego 
the use of a fiducial template necessary to predict broadband fluxes generated by dust models, and also allow direct incorporation
of dust models into our framework.

>>>>

We find weak correlations between our color-excess parameters and the input features; 
\color{orange}
only the $ -0.17 \pm 0.05$ between $E_{\gamma^0}(B-V)$ and $EW_{Si}$ has significance much larger than $1 \sigma$.
\color{black}
Recall that $EW_{Si}$ is correlated with light-curve shape, which is correlated with host-galaxy (including dust) properties 
\citep{2000AJ....120.1479H, 2003MNRAS.340.1057S}.
An extensive discussion of the correlations between the spectral features of the SNfactory data set can be found in \citet{chotard:thesis}
and \citet{leget:thesis}.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/sn2014j.pdf} 
   \caption{Top: Measured color excess for SN~2014J and the best-fit predictions from this article.   Bottom: Residual color excess
   between the data and model, with a dotted line at zero shown for reference.
   This can be compared with the best-fit results of
   \citet{2014ApJ...788L..21A} shown in their Figure~3.
   \label{sn2014j:fig}}
\end{figure}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=5.2in]{output23/deltaratio.pdf}
   \caption{Significance of the non-zero values of $\delta$ as demonstrated by the best-fit values of $\delta_X/\delta_I-1$ for $\delta$ for Model~II.
   A dotted line at zero is shown for reference.
   \label{deltaratio:fig}}
\end{figure}

@@

\color{black}
The analysis is validated using simulated datasets with and without third color parameters: the input parameters are recovered within  uncertainties.
We present here the results of one of the validation tests.  The input values of the supernova parameters are the median values from the analysis of real data.
Ten realizations of simulated data are generated based on the data covariances, and then analyzed separately.   The posteriors of the ten realizations
are aggregated and shown in Figures~\ref{simglobal1:fig}--\ref{simglobal5:fig}.  The input parameters are recovered with negligible bias
relative to the statistical uncertainties.   Note that some bias is expected: the priors for the distributions of $\Delta$, $k$, and $D$ are inconsistent
with the extracted distributions.
%There is a tendency for the slope parameters to be slightly underestimated, which is an expected consequence of 
%the uncertainties in the predictor variable, which in turn lead to 
%regression dilution.  Our analysis is susceptible to this bias because we consciously do not include distributions
%of supernova parameters as part of our model.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output23/residualx1.pdf}
    \caption{Differences between observed colors and the colors predicted from the analysis, as a function
            of the SALT2 light-curve shape parameter $x_1$.  Our model does not explicitly use $x_1$.  For reference a dotted line is plotted at zero difference.
   \label{x1res:fig}}
\end{figure}

@@

\citet{2009ApJ...699L.139W, 2011ApJ...729...55F} find a connection between $v_{Si}$ and color, and  
\citet{2015MNRAS.447.1247S} show that $v_{Si}$ is an important spectral classifier within the SNfactory data themselves.
To first order, the $\lambda_{Si}$ treated in this article varies linearly with $v_{Si}$.
The values of the $v_{Si}$ coefficient, $\eta$, in this article are consistent with zero.  $EW_{Ca}$ and $EW_{Si}$ have a significant effect on color,
and in turn $\lambda_{Si}$ is correlated with $EW_{Ca}$.
Removing from our model the dependence on equivalent widths (eliminating the  $\alpha$ and $\beta$ parameters), we recover
non-zero $\eta$ values at  $\gtrsim 2\sigma$.
The Ca and Si responsible for the equivalent widths and wavelength of our selected spectral features
are expected to be produced together in the thermonuclear burning into intermediate mass elements.
We conclude that $v_{Si}$ is correlated with color, 
but this correlation
is absorbed into the equivalent-width corrections, which give higher likelihood as they give lower residual dispersions $C_c$.

@@

Although the $D$ parameter has been referred to as ``intrinsic'', due to its placement in the model and the correspondence
of the $k$ parameters with those of extrinsic two-parameter dust models, the analysis cannot quantitatively distinguish
whether its origin is intrinsic or due to a third dust parameter or some other extrinsic effect.

@@

\subsection{Conclusion}
To summarize, we model SNe~Ia broadband optical peak magnitudes allowing for correlations with spectral features at peak and
extrinsic color parameters.  Analyzing SNfactory data with this model, we find significant evidence that the above parameters do
affect supernova magnitudes and colors.  This model  does an excellent job in
describing SN~2014J, a supernova with extreme colors that was not used in the training.  
This work represents the first determination of the dust extinction curve outside the Local Group
derived entirely independently of assumptions about the shape of the extinction curve and/or assumptions about the
distribution of $A_V$.  
\color{orange}
Nesting this model in an extended model produces a result that 
\color{black}
disfavors the original model and gives the first identification of  a new parameter that affects supernova
colors in a manner that is distinct from the expectations from dust or the spectral features $EW_{Ca}$, $EW_{Si}$, and $\lambda_{Si}$.

An interesting direction for future work is to use narrower bands for generating synthetic photometry
from the SNfactory spectrophotometry.  When the bandwidth becomes small enough
to resolve spectral features, the analysis would produce a spectroscopic model.   Higher resolution would allow us to forego 
the use of a fiducial template necessary to predict broadband fluxes generated by dust models, and also allow direct incorporation
of dust models into our framework.
